The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-06-30 Paper.Rmd
aphid count and multivar table description added

As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question 1 for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulas of all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

@@

We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
(i.e., abundance data for green spruce aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
abundance data for plankton from UK seas from $14$ sampling locations, 
methane flux data from $13$ sampling locations). We counted for how many location pair (excluding self-comparison)
copula are non-independent.

>>>>

As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question 1 for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the Normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulas of all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

@@

We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
(i.e., abundance data for green spruce aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
abundance data for plankton from UK seas from $14$ sampling locations, 
methane flux data from $13$ sampling locations). First $11$ rows in Table \ref{tab_multivar_summary} represent
model selection results :
(1) total number of location pairs (excluding self-comparison) considered, 
(2) how many location pair copula are non-independent among them, 
(3) number of non-Normal copula as best fit (based on highest AIC-weight),
(4) average AIC-weights of best fit copula from all location pair,
(5) average AIC-weights of Normal copula from all location pair,
(6) number of location pair for which best fit copula have p-values > 0.01 
(both Cramer-von Mises and Kolmogorov-Smirnov method)
(7) lower and (8) upper quantile ($95\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair,
(9) lower and (10) upper quantile ($95\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair,
(11) number of location pair which showed stronger lower tail dependence than upper.

Last $6$ rows of Table \ref{tab_multivar_summary} summarized the non-parametric analysis results 
for $\Ps$ statistics : lower and upper quantile ($95\%$CI) of $\Ps_l$, $\Ps_u$, $\Ps_l-\Ps_u$ based on resampling
location pair.

Likewise, green-spruce aphid abundance data ($1^{st}$ column in Table \ref{tab_multivar_summary}), 
plankton data from UK seas also showed major lower tail dependent spatial synchrony 
($3^{rd}$ column in Table \ref{tab_multivar_summary}), whereas firstflight data for leaf-curling plum aphid 
showed stronger upper tail dependence at every location pair ($2^{nd}$ column in Table \ref{tab_multivar_summary}).
Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
